A fast unified model for parsing and sentence understanding

### 2016-03-18 writing/hard_stack_paper/paper.tex
Trim short lines.

@@

This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}. SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly. This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences without incurring a substantial additional computational cost, removing the dependence on an external parser. In addition, it supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs. Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.

@@

The stack and the buffer can be represented as lists of $N$ elements each (for sentences of up to $N$ words), with two $D$-dimensional vectors $\vec{h}$ and $\vec{c}$ constituting each element.

@@

We use word representations based on the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}_{\text{GloVe}}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}_{\text{GloVe}} + \vec{b}_{\text{wd}}
\end{equation}

@@

\paragraph{Why a tree-sequence hybrid?} Lexical ambiguity is ubiquitous in natural language---most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence. Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which provides some cues for disambiguation. In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word. Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.

@@

We propose an alternative space-efficient stack representation technique inspired by the zipper technique \citep{huet1997zipper}. For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S_t$ represents the top of the actual stack at timestep $t$. We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time. Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation. It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run of this algorithm.

@@

\begin{table*}[t]
  \centering\small
  \begin{tabular}{lrrrr} 
    \toprule
Model                   & Params.    & Trans. acc. (\%)  &   Train acc. (\%)  &   Test acc. (\%) \\
\midrule
\multicolumn{5}{c}{\textbf{Previous non-NN results}}\\
Lexicalized classifier \citep{snli:emnlp2015}
                        & --                & --                    &   99.7   &   78.2      \\
\midrule
\multicolumn{5}{c}{\textbf{Previous sentence encoder-based NN results}}\\
100D LSTM encoders \citep{snli:emnlp2015}
                        & 221k               & --               &   84.8   &   77.6      \\
1024D pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
                        & 15m                & --              &   98.8   &   81.4       \\
300D Tree-based CNN encoders \citep{mou2015recognizing}
                        & 3.5m                & --             &   83.4   &   82.1       \\
\midrule
\multicolumn{5}{c}{\textbf{Our results}}\\
300D LSTM RNN encoders         & 3.0m                  & --                &   83.9      &   80.6       \\
300D SPINN-PI-NT (\ii{parsed input, no tracking}) encoders
                        & 3.4m                  & --                &   84.4      &   80.9       \\
300D SPINN-PI (\ii{parsed input}) encoders
                        & 3.7m                  & --                &   89.2      &   \textbf{83.2}       \\
\result{300D SPINN (unparsed input) encoders }
                        & ??                  & 9?           &   87.2    &   82.6      \\          
% \result{300D SPINN-PI, sequence-based attn.   }     
%                         & ?                  & --                &   ?      &   86.3dv         \\
% \result{300D SPINN-PI, tree-based attn. }           
%                         & ?                  & --                &   ?      &   \textbf{?}\\
% \midrule
% \multicolumn{5}{c}{Other previous NN results}\\
% \midrule
% 100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
%                         & 252k               & --              &   85.3   &   83.5       \\
% 300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
%                         & 1.9m               & --             &   92.0   &   86.1      \\
% 300D LSTMN with deep attention fusion \citep{cheng2016long}
%                         & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
    \bottomrule
  \end{tabular}
\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification. Params. is the approximate number of trained parameters (excluding word embeddings for models where they are trained). Trans. acc. is the model's accuracy in predicting parsing transitions. Train and test are SNLI classification accuracy.} 
\end{table*}

@@

We evaluate SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE;][]{dagan2006pascal,MacCartney09}. NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the SNLI corpus, which we use for training and evaluation:

\begin{description}\small 
\item[\bf Premise:] {\it Girl in a red coat, blue head wrap and jeans is making a snow angel.}
\item[\bf Hypothesis:] {\it A girl outside plays in the snow.}
\item[\bf Correct label:] entailment
\end{description}

@@

In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion. While the full models evaluated above are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here. To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}. We plan to release our modified code at publication time, alongside the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the thin-stack feedforward (not training/backpropagation) in C++/CUDA, as feedforward/inference speed is the relevant performance metric for most practical applications.}

@@

\begin{description}\small 
\item[\bf Premise:] {\it A \textbf{woman} in red blouse is \textbf{standing} with small blond \textbf{child} in front of a small folding chalkboard.}
\item[\bf Hypothesis:] {\it a woman stands with her child}
\item[\bf Correct label:] neutral
\end{description}

@@

\begin{description}\small 
\item[\bf Premise:] {\it Nurses in a medical setting conversing over a plastic cup.}
\item[\bf Hypothesis:] {\it The nurses are discussing a patient.}
\item[\bf Correct label:] neutral
\end{description}

>>>>

\newcommand{\snli}[3]{{\vspace{0.25em}
{\small \setlength{\parindent}{0.6em} \hangindent=1.2em  \textbf{Premise:} #1\par}\vspace{0.25em}
{\small \setlength{\parindent}{0.6em} \hangindent=1.2em   \textbf{Hypothesis:} #2\par}\vspace{0.25em}
{\small \setlength{\parindent}{0.6em}  \textbf{Label:} #3\par}
}}

@@

This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}. SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly. This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences without incurring a substantial additional computational cost, removing the dependence on an external parser. In addition, it supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs. Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.

@@

The stack and the buffer can be represented as arrays of $N$ elements each (for sentences of up to $N$ words), with the two $D$-dimensional vectors $\vec{h}$ and $\vec{c}$ in each element.

@@

We use word representations based on the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform each input word vector $\vec{x}_{\text{GloVe}}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}_{\text{GloVe}} + \vec{b}_{\text{wd}}
\end{equation}

@@

\paragraph{Why a tree-sequence hybrid?} Lexical ambiguity is ubiquitous in natural language---most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence. Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation. In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word. Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.

@@

We propose an alternative space-efficient stack representation technique inspired by the zipper \citep{huet1997zipper}. For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S_t$ represents the top of the actual stack at timestep $t$. We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time. Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation. It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run of this algorithm.

@@

\begin{table*}[t]
  \centering\small
  \begin{tabular}{lrrrr} 
    \toprule
Model                   & Params.    & Trans. acc. (\%)  &   Train acc. (\%)  &   Test acc. (\%) \\
\midrule
\multicolumn{5}{c}{\textbf{Previous non-NN results}}\\
Lexicalized classifier \citep{snli:emnlp2015}
                        & --                & --                    &   99.7   &   78.2      \\
\midrule
\multicolumn{5}{c}{\textbf{Previous sentence encoder-based NN results}}\\
100D LSTM encoders \citep{snli:emnlp2015}
                        & 221k               & --               &   84.8   &   77.6      \\
1024D pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
                        & 15m                & --              &   98.8   &   81.4       \\
300D Tree-based CNN encoders \citep{mou2015recognizing}
                        & 3.5m                & --             &   83.4   &   82.1       \\
\midrule
\multicolumn{5}{c}{\textbf{Our results}}\\
300D LSTM RNN encoders         & 3.0m                  & --                &   83.9      &   80.6       \\
300D SPINN-PI-NT (\ii{parsed input, no tracking}) encoders
                        & 3.4m                  & --                &   84.4      &   80.9       \\
300D SPINN-PI (\ii{parsed input}) encoders
                        & 3.7m                  & --                &   89.2      &   \textbf{83.2}       \\
300D SPINN (unparsed input) encoders
                        & 2.7m                  & \result{}           &   87.2    &   82.6      \\          
% \result{300D SPINN-PI, sequence-based attn.   }     
%                         & ?                  & --                &   ?      &   86.3dv         \\
% \result{300D SPINN-PI, tree-based attn. }           
%                         & ?                  & --                &   ?      &   \textbf{?}\\
% \midrule
% \multicolumn{5}{c}{Other previous NN results}\\
% \midrule
% 100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
%                         & 252k               & --              &   85.3   &   83.5       \\
% 300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
%                         & 1.9m               & --             &   92.0   &   86.1      \\
% 300D LSTMN with deep attention fusion \citep{cheng2016long}
%                         & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
    \bottomrule
  \end{tabular}
\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification. Params. is the approximate number of trained parameters (excluding word embeddings for models where they are trained). Trans. acc. is the model's accuracy in predicting parsing transitions. Train and test are SNLI classification accuracy.} 
\end{table*}

@@

We evaluate SPINN on the task of natural language inference \citep[NLI, also known as recognizing textual entailment, or RTE;][]{dagan2006pascal,MacCartney09}. NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the SNLI corpus, which we use:

\snli{Girl in a red coat, blue head wrap and jeans is making a snow angel.}
{A girl outside plays in the snow.}
{entailment}

@@

In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion. While the full models evaluated above are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here. To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}. We plan to release our modified code at publication time, alongside the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference in C++/CUDA, as inference speed is the relevant performance metric for most practical applications.}

@@

\snli
{A \textit{woman} in red blouse is \textit{standing} with small blond \textit{child} in front of a small folding chalkboard.}
{a woman stands with her child}
{neutral}

@@

\snli{Nurses in a medical setting conversing over a plastic cup.}
{The nurses are discussing a patient.}
{neutral}
